\blue{Numerical Investigation of Multi-Fluid Flows through Dual Saturated-Unsaturated Porous Media - change the title to something close to the main topic of the manuscript ... something with viscous flow instability or viscous fingering)}

### 2017-03-02 Papers/ViscousInstabilities/Manuscript_ViscousInstabilities_KC.tex
updated

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\section{Numerical Formulation}\label{section:NumericalFormulation}
 The method developed is applicable to $N_{p}$ fluid phases and is based on the novel family of $P_{n}(DG)-P_{m}(DG)$ element-pair consistent with the dual pressure-velocity representation in control volume (CV) and finite element (FE) spaces. These element-pairs have a discontinuous $n^{th}$-order polynomial representation for velocity and a discontinuous $m^{th}$-order polynomial pressure. Another element pair, that can be used is the $P_{n}DGP_{m}$ with an $m^{th}$-order continuous representation for pressure. The mass balance (continuity) equations are solved in CV space and a Petrov-Galerkin finite element (FE) method is used to obtain the high-order fluxes on CV boundaries (fig.\ref{fig:fem_elem})  , which are limited to yield bounded fields.

@@

\medskip
This viscous driven hydrodynamic instability results the collapse of the uniform interface between the fluids of different viscosity. This type of instability can be found across many research areas, disciplines and scales, from small chemical and medical separation systems to large scale field geological and /or reservoir systems. Starting from the work of \citet{saffman_1959} and \citet{homsy_1987}, and the experimental setup of Hele-Shaw cell by \citet{saffman_1986}, a basic mathematical model has been initially developed to describe the physics of this phenomenon.

\medskip
In viscous fingering, a less viscous fluid (i.e. water) pushes a more viscous fluid (i.e. oil), in a thin rectangular cells (known as Hele-Shaw cells). Under the assumption that fluids remaining completely separated (or immiscible) along an interface, surface tension plays an important role in determining the shape and progress of the fingers. For homogeneous domains fingers will start to develop when surface tension acting on the interface, is opposing towards the change of shape of the interface thus the interface becomes unstable and will collapse, taking a curved shape. In heterogeneous environments this process may also be triggered by permeability differences between smaller regions. The higher the velocity of the injected fluid, the less wide (a tip-splitting behaviour) the finger will be. In addition, pressure differences (eq.\ref{eq:pressure_dif}) acting on the interface will produce a net pressure force.

\begin{equation}
\Delta P= - \gamma \nabla \; \dot \; \hat{n} 
\label{eq:pressure_dif} 
\end{equation}

\noindent This is also know as the Young-Laplace equation, a relation describing the capillary pressure difference across the interface between two fluids. With $\Delta P$ denoting the pressure difference,$\hat{n}$ the unit normal vector out of the surface, and $\gamma$ the surface tension. As mentioned in section \ref{section:intro}, an unfavourable mobility-ratio (i.e. in a water-oil system, see eq.\ref{eq:MR}, displacement coupled with the heterogeneity of the porous medium results in viscous instability, which in turn affects the displacement efficiency of im/miscible processes.

@@

\noindent Fingering will occur in miscible displacements when the viscocity ratio is greater than unity. %As surface tension becomes weak, the front of the steady finger is susceptible to a viscous-fingering instability by the basic mechanism associated with a less viscous fluid displacing a more viscous one. After a split, each of the new lobes of the finger is stable as a result of their being thinner than the finger from which they split. 
As a result of \textit{shielding}, one of these fingers will eventually outgrow the others and will then spread to occupy the appropriate width of the cell. In the process, it reaches a width that is again unstable to splitting, and the pattern repeats. Thus, surface tension plays an essential dual role; it must be weak enough for the tip front to be unstable, but it is also the physical force causing the \textit{spreading} and ensuing repeated branching.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\section{Numerical Formulation}\label{section:NumericalFormulation}
 The method developed is applicable to $N_{p}$ fluid phases and is based on the novel family of $P_{n}(DG)-P_{m}(DG)$ element-pair consistent with the dual pressure-velocity representation in control volume (CV) and finite element (FE) spaces. These element-pairs have a discontinuous $n^{th}$-order polynomial representation for velocity and a discontinuous $m^{th}$-order polynomial pressure. Another element pair, that can be used is the $P_{n}DGP_{m}$ with an $m^{th}$-order continuous representation for pressure. The mass balance (continuity) equations are solved in CV space and a Petrov-Galerkin finite element (FE) method is used to obtain the high-order fluxes on CV boundaries (fig.\ref{fig:fem_elem}), which are limited to yield bounded fields.

@@

\medskip
This viscous driven hydrodynamic instability leads to the collapse of the uniform interface between fluids of different viscosity. This type of instability can be found across many research areas, disciplines and scales, from small chemical and medical separation systems to large scale field geological and /or reservoir systems. Starting from the work of \citet{muskat_1934} and later from \citet{saffman_1959} and \citet{homsy_1987}, the experimental setup of Hele-Shaw cell by \citet{saffman_1986}, and the basic mathematical model has been initially developed to describe the physics of this phenomenon. The idea is to introduce a fluid between two parallel flat plates separated by an infinitesimally small gap. The governing equation of Hele-Shaw flows is identical to that of the inviscid potential flow and to the flow of fluid through a porous medium (Darcy's law). More recently work that has been done by \citet{howison_2000} and \citet{praud_2005} give an extensive description on the Saffman $\&$ Taylor problem (also know as immiscible two-phase flow in Hele-Shaw cell).

\medskip
In viscous fingering, a less viscous fluid pushes a more viscous fluid, in a thin rectangular cell (known as Hele-Shaw cell). \citet{howison_2000}  Under the assumption that fluids remain immiscible along an interface, surface tension plays an important role in determining the shape and progress of the fingers. For homogeneous domains fingers will start to develop when surface tension acting on the interface, is opposing towards the change of shape of the interface thus the interface becomes unstable and will collapse, taking a curved shape. In heterogeneous environments this process may also be triggered by permeability differences between smaller regions. The higher the velocity of the injected fluid, the less wide (tip-splitting behaviour) the finger will be. In addition, pressure differences %(eq.\ref{eq:pressure_dif}) 
acting on the interface will produce a net pressure force,

\begin{equation}
\Delta P= - \gamma \; \nabla \; \dot \; \hat{n} 
\label{eq:pressure_dif} 
\end{equation}

\noindent this is also know as the Young-Laplace equation, a relation describing the capillary pressure difference across the interface between two fluids. With $\Delta P$ denoting the pressure difference, $\hat{n}$ the unit normal vector out of the surface, and $\gamma$ the surface tension. As mentioned in Section \ref{section:intro}, an unfavourable mobility-ratio in a water-oil system will have the following definition, see eqn.\ref{eq:MR}.

@@

\noindent Fingering will occur in immiscible displacements when the viscocity ratio is greater than unity. %As surface tension becomes weak, the front of the steady finger is susceptible to a viscous-fingering instability by the basic mechanism associated with a less viscous fluid displacing a more viscous one. After a split, each of the new lobes of the finger is stable as a result of their being thinner than the finger from which they split. 
As a result of \textit{shielding}, one of these fingers will eventually outgrow the others and will then spread to occupy the appropriate width of the cell. In the process, it reaches a width that is again unstable to splitting, and the pattern repeats. Thus, surface tension plays an essential dual role; it must be weak enough for the tip front to be unstable, but it is also the physical force causing the \textit{spreading} and ensuing repeated branching.
